What paradigms can webcam eye-tracking be used for? Attempted replications of five ``classic'' cognitive science experiments

### 2025-04-02 manuscript.Rmd
minor changes

---
title             : "What paradigms can webcam eye-tracking be used for? Attempted replications of five cognitive science experiments"
shorttitle        : "Webcam eye-tracking paradigms"
author:
  - name: Ariel N. James
    affiliation: '1'
    role:
      - Conceptualization
      - Formal analysis
      - Visualization
      - Writing - original draft
      - Writing - review & editing
    email: ajames2@macalester.edu
    corresponding: yes
    address: 1600 Grand Ave, Saint Paul, MN 55105, USA

@@

  Web-based data collection allows researchers to recruit large and diverse samples with fewer resources than lab-based studies require. Recent innovations have expanded the set of methodolgies that are possible online, but ongoing work is needed to test the suitability of web-based tools for various research paradigms. Here, we focus on webcam-based eye-tracking; we tested whether the results of five different eye-tracking experiments in the cognitive psychology literature would replicate in a webcam-based format. Specifically, we carried out five experiments by integrating two javascript-based tools: jsPsych and a modified version of Webgazer.js. In order to represent a wide range of applications of eye-tracking to cognitive psychology, we chose two psycholinguistic experiments, two memory experiments, and a decision-making experiment. These studies also varied in the type of eye-tracking display, including screens split into halves (Exps. 3 and 5) or quadrants (Exps. 2 and 4), or composed scenes with regions of interest that varied in size (Exp. 1). Outcomes were mixed. The least successful replication attempt was Exp. 1; we did not obtain a condition effect in our remote sample (1a), nor in an in-lab follow-up (1b). However, the other four experiments were more successful, replicating a blank-screen effect (Exp. 2), a novelty preference (Exp. 3), a verb bias effect (Exp. 4), and a gaze-bias effect in decision-making (Exp. 5). These results suggest that webcam-based eye tracking can be used to detect a variety of cognitive phenomena, including those with sensitive time, although paradigms that require high spatial resolution (like Exp. 1) should be adapted to coarser quadrant or split-half displays. 
keywords          : "eye-tracking, online, webcam, jsPsych, cognitive science"
wordcount         : "13,115"
bibliography      : ["references.bib"]
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

@@

The use of eye-tracking to study cognition took off when Alfred Yarbus
used suction cups to affix a mirror system to the sclera of the eye in
order to monitor eye position during the perception of images
[@yarbus1967eye]. In one study, participants viewed a painting depicting
multiple people in a complex interaction inside of a 19th century
Russian home. Yarbus showed, among other things, that the scan paths and
locations of fixations were largely dependent on the instructions given
to participants (e.g., View the picture freely vs. Remember the position
of the people and objects in the room). In other words, the cognitive
processing that the individual is engaged in drives the visuo-motor
system. Since these findings, eye-tracking has become a central method
in cognitive science research [for reviews, see @hayhoe2005eye,
@richardsonEyeTrackingResearch2004, @rayner1998eye]. For example, gaze
location during natural scene perception is used to test theories of
visual attention [e.g., @henderson2017meaning], and eye-movements during
auditory language comprehension, using the "visual world paradigm,"
demonstrated the context-dependent and incremental nature of language
processing [e.g., @tanenhaus1995].

An important limitation of the eye-tracking methodology is that it has
typically required costly equipment (eye-trackers can range in price
from a few thousand dollars to tens of thousands of dollars), particular
laboratory conditions (a quiet room with consistent indoor lighting
conditions), and a substantial time investment (e.g., bringing
participants into a laboratory one at a time). This limits who can
conduct eye-tracking research -- not all researchers have the necessary
resources -- and who can participate in eye-tracking research. Most
eye-tracking study participants are from western, educated,
industrialized, rich, and democratic [WEIRD; @henrich2010weirdest]
convenience samples [but see @ryskin2023real], which diminishes the
generalizability of the findings and the scope of conclusions that can
be drawn about human cognition. Likewise, the sample sizes for in-lab
experiments are usually orders of magnitude smaller than what
statisticians recommend [@nosek2022replicability].

A robust solution to all these problems is online experiments,
particularly with volunteer citizen scientists as participants
[@li2024developmental; @hartshorne2019thousand;
@reinecke2015labinthewild; @gosling2010wired]. Historically, this option
has not been available for eye-tracking, since few potential subjects
have expensive eyetrackers at home. In principal, this could be
circumvented by using so-called "poor-man's" eye-tracking, in which
researchers videotape the subject's face and code eye position
frame-by-frame [@snedeker2003using]. Most computers now have
high-quality built-in cameras, so recording video is straightforward.
Unfortunately, coding images frame-by-frame is extremely time-intensive,
making large-sample studies unrealistic. However, in recent years, image
analysis has improved to the point where this work can be automated with
reasonable accuracy [@papoutsaki2016webgazer; @zheng2018rapid;
@burton2014comparison; @skovsgaard2011evaluation]. Nonetheless,
webcam-based eyetracking only started to be used regularly in research
with the advent of `Webgazer.js` [@papoutsaki2016webgazer], a
webcam-based Javascript plug-in that works in the browser and which can
be integrated with any Javascript web interface, including `jsPsych`
[@deleeuwJsPsychJavaScriptLibrary2015], `Gorilla` [@anwyl2020gorilla],
or `lab.js` [@henninger2021lab].

@@

While promising, there are some salient limitations. First, many of the
studies report effects that are smaller or later than what had been
previously observed in the lab [@slim2022moving; @degen2021seeing;
@van2023validation]. Importantly, accurate timing on a web browser is
not trivial [@passell2021cognitive; @de2016psychophysics], and subtle
programming choices can significantly affect the accuracy of
`WebGazer.js` timing [@yang2021webcam].[^1] Since most of the prior work
did not address these timing issues, it is not clear how many of the
reported lab/web differences would resolve.

[^1]: See also discussion at
    <https://github.com/jspsych/jsPsych/discussions/1892>

Second, prior work has focused on studies with relatively coarse-grained
regions of interest [but see @semmelmann2018online], dividing the screen
either in half or in quadrants. This is particularly salient with
[@prystauka2023online], which simplified
@altmannIncrementalInterpretationVerbs1999's design so that regions of
interest are quadrants rather than the finer-grained ROIs used in the
original. Certainly, webcam eyetracking will not be as spatially
fine-grained as an infrared eyetracker, but we do not yet have a good
sense of the limits.

@@

<!-- June 2022: shimojo: 964 citations, altmann: 1840, ryskin: 75, johansson 190, 127 -->

<!-- RR: switched ryskin to Snedeker & trueswell 2004 bcs my 75 looked sad -->

We chose five high-impact eye-tracking studies involving adult subjects
[for an comparison of remote WebGazer and in-lab anticipatory looking
effects with 18-27 month-old participants, see @steffan2024validation].
Our goal was to include experiments from a range of topic areas (e.g.,
memory, decision making, psycholinguistics) and paradigms (two halves of
the screen, visual world paradigm with four quadrants, visual world
paradigm with "naturalistic" scenes). As noted above, we had a
preference for well-established findings that are known to replicate,
though for sake of diversity this was not always possible. Table
\@ref(tab:studies-table) provides an overview of the five studies we
selected.

@@

These data were collected within the context of an undergraduate
research methods course. Groups of students (co-authors) designed and
programmed experiments in jsPsych, pre-registered their planned
analyses, and collected data through Prolific under the supervision of
the last author. The OSF repositories associated with these experiments
are linked in the methods sections of each individual study. Note that
in the current paper we expand on those pre-registered analyses (e.g.,
including analyses of the calibration quality). All analysis code
underlying this paper can be found in the Github repository:
<https://github.com/jodeleeuw/219-2021-eyetracking-analysis>

@@

The first study was a replication attempt of
@altmannIncrementalInterpretationVerbs1999. Altmann and Kamide used the
visual world eye-tracking paradigm [@tanenhaus1995] to show that
meanings of verbs rapidly constrain the set of potential subsequent
referents in sentence processing. For example, when looking at the
display in Figure \@ref(fig:E1-example-trial) and listening to a
sentence like "The boy will eat the...," participants are more likely to
look at the cake than when they hear "The boy will move the...," in
which case they tend to look at the train, presumably because cakes are
edible and trains are not. Semantic information available at the verb is
used to anticipate upcoming linguistic input.

@@

Sixty participants (2.5 times the original sample of 24) were paid
\$2.60 for their participation. For unknown reasons, 2 of the subjects'
results were not recorded, so in the analysis, we worked with data
collected from 58 participants.

@@

The visual stimuli were created through Canva and depicted an agent
accompanied by four to five objects in the scene (see
Figure \@ref(fig:E1-example-trial)). On critical trials, participants
heard one of two sentences associated with the scene. In the restrictive
condition, the sentence (e.g., "The boy will eat the cake") contained a
verb (e.g., "eat") which restricts the set of possible subsequent
referents (e.g., to edible things). Only the target object (e.g., the
cake) was semantically consistent with the verb's meaning. In the
non-restrictive condition, the sentence (e.g., "The boy will move the
cake") contained a verb (e.g., "move") which does not restrict the set
of possible subsequent referents. The target object (e.g., the cake) as
well as the distractor objects (e.g., the train, the ball, etc.) were
semantically consistent with the verb's meaning. Both sentences were
compatible with the scene, such that the correct keyboard response for
the critical trials was "yes." Filler trials consisted of scenes that
also contained an agent surrounded by objects as in the critical trials,
but corresponding sentences named an object that was not present in the
scene. The correct keyboard response for the filler trials was "no."

@@

```{=html}
<!-- Qs: 
-keyboard keys? what was the exact phrasing of the instruction? 
-Any delay between when visual scene appears and the audio starts? 
-Does trial end as soon as participant responds?-->
```
## Results

Looks to the objects in the scene were time-locked to the onset of the
verb, the offset of the verb, onset of the post-verbal determiner, and
onset of the target noun. ROIs were defined by creating boxes around
each object in the scene. The size of each box was determined by taking
the height and width of the given object and adding 20 pixels of
padding. Each scene contained an agent region, a target region, and
three or four distractor regions.

@@

Across three different tests of the hypothesis that listeners will use
verb semantics to anticipate the upcoming referent, we found results
that were numerically consistent with the hypothesis but not
statistically significant. This is unlikely to be because the original
effect is unreliable. Indeed, one of us had successfully replicated it
previously using traditional methods [@james2023language].

@@

# Experiment 1b

Experiment 1b tested whether the failure to replicate significant
condition effects in Experiment 1a was due to features of conducting the
study remotely (i.e. varied experimental settings and apparatuses, lower
compliance) rather than webcam-based eye-tracking or `webgazer` *per
se*. Thus, Experiment 1b took place in a lab setting with undergraduate
participants but otherwise used the same Method as Experiment 1a.

@@

Forty-nine participants completed the study in a lab setting. They were
recruited via the Boston College subject pool. Participants needed to be
18 years of age or older and native speakers of English.

Materials were identical to those in Experiment 1a.

After being greeted by the experimenter and completing the informed
consent form, participants followed on-screen prompts to complete the
study, including calibration, as described in the Experiment 1a
Procedure.

### Apparatus

Subjects were tested on a MacIntosh laptop using the built-in Webcam.
The study was run in Chrome using the same code as Exp. 1a.

```{r E1-analysis-JoshH-lab, child = 'group-a/E1-analysis-comparison-to-Hartshorne.Rmd'}

@@

Overall, the results of Experiment 1 paint a sobering picture of
web-based eye-tracking. In Experiment 1a, results from the remote sample
were in the expected direction but effects were smaller and delayed
relative to previous work and failed to reach statistical significance.
This was not due to vagaries of remote samples or working with a more
diverse subject population: results from the in-lab study were, if
anything, worse. Nor does our failure to replicate seem to be driven by
subjects with poor calibration; effect sizes were not correlated with
calibration success, and restricting analyses to subjects with good
calibration scores only helped in one of the experiments and in only one
of the three critical analyses.

@@

The experiment consisted of two blocks each composed of an encoding
phase and a recall phase. The two blocks differed by whether the recall
phase was in the free-viewing or fixed-viewing condition, as described
in the Procedure. Participants were randomly assigned to see the
fixed-viewing or free-viewing block first. During each encoding phase,
participants saw a grid indicating the four quadrants of the screen.
Each quadrant contained six cartoon images of items belonging to the
same category. The four categories were humanoids, household objects,
animals, and methods of transportation (see
Figure \@ref(fig:E2-example-trial)). Different images were used in each
block; there were 48 unique images total across the experiment.

@@

Stimulus sets were created to present participants with 24 presentation
trials and 24 test trials on Day 1, and 48 recognition trials on Day 2.
Each presentation trial screen presented two identical images (e.g. red
apple and red apple). The test trial screens presented each of the
previously-seen ("old") images paired with a new image (e.g. red apple
and bicycle key). The recognition trial screens presented a single image
that was either from the original set of 24 (seen twice across the first
two phases) or was a corresponding foil for the original set (e.g. red
apple or green apple). Thus, each participant was exposed to 72 unique
images over the course of the experiment (the 24 foils for the "new"
images in the test phase are never seen). Two full stimulus lists were
created to counterbalance images across participants; the images that
composed the "old" set for one list made up the "new" set for the list.
The experimental design is visually depicted in
Figure \@ref(fig:E3-design-schematic).

@@

During a trial, four pictures were displayed (target animal, target
instrument, distractor animal, distractor instrument), one in each
corner of the screen, and participants heard an audio prompt that
contained instructions about the action they needed to act out (e.g.,
"Rub the butterfly with the crayon"; see
Figure \@ref(fig:E4-example-trial))[^2]. Using their cursor,
participants could act out the instructions by clicking on objects and
moving them or motioning over the objects[^3]. After the action was
completed, the participants were instructed to press the space bar which
led to a screen that said "Click Here" in the middle in order to remove
bias in the eye and mouse movements from the previous trial. The
experiment only allowed the participants to move on to the next trial
once the audio was completely done playing and the mouse had been moved
over at least one object.

@@

All stimuli, experiment scripts, data, and analysis scripts are
available on the Open Science Framework at <https://osf.io/eubsc/>. The
study pre-registration is available at <https://osf.io/tv57s>.

@@

During each trial of the main task, two faces were displayed on the two
halves of the screen, one on the left and one on the right (as in
Figure \@ref(fig:E5-example-trial)). In the attractiveness task,
participants were asked to chose the more attractice face in the pair
and in the shape judgment task participants were asked to pick the face
that appeared rounder. They pressed the "a" key on their keyboard to
select the face on the left and the "d" key to select the face on the
right. A fixation cross appeared in the center of the screen between
each set of faces. Participants were asked to look at this fixation
cross in order to reset their gaze in between trials. The order of the
19 face pairs was random for each participant.

@@

In the original study, a video-based eye tracker was used. The eye
movements of participants were recorded with a digital camera
downsampled to 33.3 Hz, with eye position determined automatically with
MediaAnalyzer software. In our study, subjects supplied their own
cameras, so hardware sampling rate varied. However, data was collected
at 20 Hz.
<!--`r colorize("[TODO - CONFIRM]", 'red')` AJ: I don't know how to confirm this! -->

@@

 -->
```
# General Discussion

@@

Note also that while ROIs the size of quadrants are feasible, good
calibration is important. Based on the results above, researchers using
such paradigms may want to exclude data from participants with less than
75% validation accuracy, whereas studies using two halves of the display
as ROIs may not need to be so conservative.

Second, researchers need to take care with timing accuracy. Running
studies on modern computers requires synchronizing multiple clocks,
something that does not happen automatically. Failure to synchronize can
result in substantial inaccuracies [@slim2022moving]. As indicated by
the current work, the current jsPsych implementation of WebGazer.js has
reasonably good timing synchronization. We recommend that future
researchers use jsPsych or another package with similar documented
precision.

Third, it is true for any eyetracker that precise data depends on good
callibration, and good callibration depends in part on lighting and
background. We suspect that it is possible to give subjects automated
feedback on how to improve their own calibration, which could result in
higher-quality data and fewer subject exclusions based on poor
calibration.

@@

# Contributions

-   Joshua R. de Leeuw
    -   Conceptualization
    -   Data curation
    -   Formal analysis
    -   Investigation
    -   Methodology
    -   Project administration
    -   Software
    -   Supervision
    -   Validation
    -   Visualization
    -   Writing - original draft
    -   Writing - review & editing
-   Ariel N. James, Rachel Ryskin, and Joshua K. Hartshorne
    -   Conceptualization
    -   Formal analysis
    -   Visualization
    -   Writing - original draft
    -   Writing - review & editing
-   Undergraduate student authors, listed alphabetically: Haylee Backs,
    Nandeeta Bala, Laila Barcenas-Meade, Samata Bhattarai, Tessa
    Charles, Gerasimos Copoulos, Claire Coss, Alexander Eisert, Elena
    Furuhashi, Keara Ginell, Anna Guttman-McCabe, Emma (Chaz) Harrison,
    Laura Hoban, William A. Hwang, Claire Iannetta, Kristen M. Koenig,
    Chauncey Lo, Victoria Palone, Gina Pepitone, Margaret Ritzau, Yi Hua
    Sung, and Lauren Thompson
    -   Investigation
    -   Methodology
    -   Software

@@

# Funding information

Funding for data collection was provided by the Vassar College
Department of Cognitive Science (JdL). Partial funding for JKH's
contributions was provided by NSF2436092.

@@

# Competing interests

The authors have no competing interests to declare.

@@

# Data accessibility statement

All data and analysis code underlying this paper can be found in the
Github repository:
<https://github.com/jodeleeuw/219-2021-eyetracking-analysis>. Study
materials and pre-registrations for each experiment are available on
Open Science Framework:

>>>>

---
title             : "What paradigms can webcam eye-tracking be used for? Attempted replications of five cognitive science experiments"
shorttitle        : "Webcam eye-tracking paradigms"

author:
  - name: Ariel N. James
    affiliation: '1'
    role:
      - Conceptualization
      - Formal analysis
      - Visualization
      - Writing - original draft
      - Writing - review & editing
    email: ajames2@macalester.edu
    corresponding: yes
    address: 1600 Grand Ave, Saint Paul, MN 55105, USA

@@

  Web-based data collection allows researchers to recruit large and diverse samples with fewer resources than lab-based studies require. Recent innovations have expanded the set of methodolgies that are possible online, but ongoing work is needed to test the suitability of web-based tools for various research paradigms. Here, we focus on webcam-based eye-tracking; we tested whether the results of five different eye-tracking experiments in the cognitive psychology literature would replicate in a webcam-based format. Specifically, we carried out five experiments by integrating two javascript-based tools: jsPsych and a modified version of Webgazer.js. In order to represent a wide range of applications of eye-tracking to cognitive psychology, we chose two psycholinguistic experiments, two memory experiments, and a decision-making experiment. These studies also varied in the type of eye-tracking display, including screens split into halves (Exps. 3 and 5) or quadrants (Exps. 2 and 4), or composed scenes with regions of interest that varied in size (Exp. 1). Outcomes were mixed. The least successful replication attempt was Exp. 1; we did not obtain a condition effect in our remote sample (1a), nor in an in-lab follow-up (1b). However, the other four experiments were more successful, replicating a blank-screen effect (Exp. 2), a novelty preference (Exp. 3), a verb bias effect (Exp. 4), and a gaze-bias effect in decision-making (Exp. 5). These results suggest that webcam-based eye tracking can be used to detect a variety of cognitive phenomena, including those with sensitive time, although paradigms that require high spatial resolution (like Exp. 1) should be adapted to coarser quadrant or split-half displays.  
keywords          : "eye-tracking, online, webcam, jsPsych, cognitive science"
wordcount         : "X"
bibliography      : ["references.bib"]
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

@@

The use of eye-tracking to study cognition took off when Alfred Yarbus used suction cups to affix a mirror system to the sclera of the eye in order to monitor eye position during the perception of images [@yarbus1967eye]. In one study, participants viewed a painting depicting multiple people in a complex interaction inside of a 19th century Russian home. Yarbus showed, among other things, that the scan paths and locations of fixations were largely dependent on the instructions given to participants (e.g., View the picture freely vs. Remember the position of the people and objects in the room). In other words, the cognitive processing that the individual is engaged in drives the visuo-motor system. Since these findings, eye-tracking has become a central method in cognitive science research [for reviews see @hayhoe2005eye; @richardsonEyeTrackingResearch2004; @rayner1998eye]. For example, gaze location during natural scene perception is used to test theories of visual attention [e.g., @henderson2017meaning], and eye-movements during auditory language comprehension, using the "visual world paradigm," demonstrated the context-dependent and incremental nature of language processing [e.g., @tanenhaus1995].

An important limitation of the eye-tracking methodology is that it has typically required costly equipment (eye-trackers can range in price from a few thousand dollars to tens of thousands of dollars), particular laboratory conditions (a quiet room with consistent indoor lighting conditions), and a substantial time investment (e.g., bringing participants into a laboratory one at a time). This limits who can conduct eye-tracking research -- not all researchers have the necessary resources -- and who can participate in eye-tracking research. Most eye-tracking study participants are from western, educated, industrialized, rich, and democratic [WEIRD, @henrich2010weirdest] convenience samples [but see @ryskin2023real], which diminishes the generalizability of the findings and the scope of conclusions that can be drawn about human cognition. Likewise, the sample sizes for in-lab experiments are usually orders of magnitude smaller than what statisticians recommend [@nosek2022replicability].

A robust solution to all these problems is online experiments, particularly with volunteer citizen scientists as participants [@li2024developmental; @hartshorne2019thousand; @reinecke2015labinthewild; @gosling2010wired]. Historically, this option has not been available for eye-tracking, since few potential subjects have expensive eyetrackers at home. In principal, this could be circumvented by using so-called "poor-man's" eye-tracking, in which researchers videotape the subject's face and code eye position frame-by-frame [@snedeker2003using]. Most computers now have high-quality built-in cameras, so recording video is straightforward. Unfortunately, coding images frame-by-frame is extremely time-intensive, making large-sample studies unrealistic. However, in recent years, image analysis has improved to the point where this work can be automated with reasonable accuracy [@papoutsaki2016webgazer; @zheng2018rapid; @burton2014comparison; @skovsgaard2011evaluation]. Nonetheless, webcam-based eyetracking only started to be used regularly in research with the advent of `Webgazer.js` [@papoutsaki2016webgazer], a webcam-based Javascript plug-in that works in the browser and which can be integrated with any Javascript web interface, including `jsPsych` [@deleeuwJsPsychJavaScriptLibrary2015], `Gorilla` [@anwyl2020gorilla], `lab.js` [@henninger2021lab], or `PsychoJS` [@peirce2019psychopy2].

@@

While promising, there are some salient limitations. First, many of the studies report effects that are smaller or later than what had been previously observed in the lab [@slim2024webcams; @van2024validation; @slim2022moving; @degen2021seeing; @van2023validation; @bogdan2024investigating; @kandel2024assessing]. Some of this is likely due to programming errors rather than inherent limitations in the technology: Importantly, accurate timing on a web browser is not trivial [@passell2021cognitive; @de2016psychophysics], and subtle programming choices can significantly affect the accuracy of `WebGazer.js` timing [@yang2021webcam].^[See also discussion at https://github.com/jspsych/jsPsych/discussions/1892] Since most of the prior work did not address these timing issues, it is not clear how many of the reported lab/web differences would resolve.

Second, prior work has focused on studies with relatively coarse-grained regions of interest [but see @semmelmann2018online; @bogdan2024investigating], dividing the screen either in half or in quadrants. This is particularly salient with [@prystauka2023online], which simplified @altmannIncrementalInterpretationVerbs1999's design so that regions of interest are quadrants rather than the finer-grained ROIs used in the original. Certainly, webcam eyetracking will not be as spatially fine-grained as an infrared eyetracker, but we do not yet have a good sense of the limits.

@@

We chose five high-impact eye-tracking studies involving adult subjects [for a comparison of remote WebGazer and in-lab anticipatory looking effects with 18-27 month-old participants, see @steffan2024validation]. Our goal was to include experiments from a range of topic areas (e.g., memory, decision making, psycholinguistics) and paradigms (two halves of the screen, visual world paradigm with four quadrants, visual world paradigm with "naturalistic" scenes). As noted above, we had a preference for well-established findings that are known to replicate, though for sake of diversity this was not always possible. Table \@ref(tab:studies-table) provides an overview of the five studies we selected.

@@

These data were collected within the context of an undergraduate research methods course. Groups of students (co-authors) designed and programmed experiments in jsPsych, pre-registered their planned analyses, and collected data through Prolific under the supervision of the last author. The OSF repositories associated with these experiments are linked in the methods sections of each individual study. Note that in the current paper we expand on those pre-registered analyses (e.g., including analyses of the calibration quality).

@@

The first study was a replication attempt of
@altmannIncrementalInterpretationVerbs1999. Altmann and Kamide used the
visual world eye-tracking paradigm [@tanenhaus1995] to show that
meanings of verbs rapidly constrain the set of potential subsequent
referents in sentence processing. For example, when looking at the
display in Figure\ \@ref(fig:E1-example-trial) and listening to a sentence like
"The boy will eat the...," participants are more likely to look at the cake than 
when they hear "The boy will move the...," in which case they tend to look at the
train, presumably because cakes are edible and trains are not. Semantic
information available at the verb is used to anticipate upcoming
linguistic input.

@@

### Participants
Sixty participants (2.5 times the original sample of 24) were paid \$2.60 for their participation. For unknown reasons,
2 of the subjects' results were not recorded, so in the analysis, we
worked with data collected from 58 participants.

@@

The visual stimuli were created through Canva and depicted an agent
accompanied by four to five objects in the scene (see Figure\ \@ref(fig:E1-example-trial)). On
critical trials, participants heard one of two sentences associated with
the scene. In the restrictive condition, the sentence (e.g., "The boy
will eat the cake") contained a verb (e.g., "eat") which restricts the
set of possible subsequent referents (e.g., to edible things). Only the
target object (e.g., the cake) was semantically consistent with the
verb's meaning. In the non-restrictive condition, the sentence (e.g.,
"The boy will move the cake") contained a verb (e.g., "move") which does
not restrict the set of possible subsequent referents. The target object
(e.g., the cake) as well as the distractor objects (e.g., the train, the
ball, etc.) were semantically consistent with the verb's meaning. Both
sentences were compatible with the scene, such that the correct keyboard
response for the critical trials was "yes." Filler trials consisted of
scenes that also contained an agent surrounded by objects as in the critical trials, but corresponding sentences named an object that was not present in the scene. The correct keyboard response for the filler
trials was "no."

@@

```{=html}
<!-- Qs: 
-keyboard keys? what was the exact phrasing of the instruction? 
-Any delay between when visual scene appears and the audio starts? 
-Does trial end as soon as participant responds?-->
```

## Results
Looks to the objects in the scene were time-locked to the onset of the
verb, the offset of the verb, onset of the post-verbal determiner, and
onset of the target noun. ROIs were defined by creating boxes around each object in the scene. The size of each box was determined by taking the height and width of the given object and adding 20 pixels of padding. Each scene contained an agent region, a target region, and three or four distractor regions.

@@

## Discussion
Across three different tests of the hypothesis that listeners will use verb semantics to anticipate the upcoming referent, we found results that were numerically consistent with the hypothesis but not statistically significant. This is unlikely to be because the original effect is unreliable. Indeed, one of us had successfully replicated it previously using traditional methods [@james2023language].

@@

# Experiment 1b
Experiment 1b tested whether the failure to replicate significant condition effects in Experiment 1a was due to features of conducting the study remotely (i.e. varied experimental settings and apparatuses, lower compliance) rather than webcam-based eye-tracking or `webgazer` _per se_. Thus, Experiment 1b took place in a lab setting with undergraduate participants but otherwise used the same Method as Experiment 1a.

@@

### Participants
Forty-nine participants completed the study in a lab setting. They were recruited via the Boston College subject pool. Participants needed to be 18 years of age or older and native speakers of English.

### Materials and Design
Materials were identical to those in Experiment 1a.

### Procedure
After being greeted by the experimenter and completing the informed consent form, participants followed on-screen prompts to complete the study, including calibration, as described in the Experiment 1a Procedure.

### Apparatus
Subjects were tested on a MacIntosh laptop using the built-in Webcam. The study was run in Chrome using the same code as Exp. 1a.

## Results
```{r E1-analysis-JoshH-lab, child = 'group-a/E1-analysis-comparison-to-Hartshorne.Rmd'}

@@

## Discussion
Overall, the results of Experiment 1 paint a sobering picture of web-based eye-tracking. In Experiment 1a, results from the remote sample were in the expected direction but effects were smaller and delayed relative to previous work and failed to reach statistical significance. This was not due to vagaries of remote samples or working with a more diverse subject population: results from the in-lab study were, if anything, worse. Nor does our failure to replicate seem to be driven by subjects with poor calibration; effect sizes were not correlated with calibration success, and restricting analyses to subjects with good calibration scores only helped in one of the experiments and in only one of the three critical analyses.

@@

The experiment consisted of two blocks each composed of an encoding
phase and a recall phase. The two blocks differed by whether the recall phase was in the free-viewing or fixed-viewing condition, as described in the Procedure. Participants were randomly assigned to see the
fixed-viewing or free-viewing block first. 
During each encoding phase, participants saw a
grid indicating the four quadrants of the screen. Each quadrant
contained six cartoon images of items belonging to the same category. The four categories were humanoids, household objects, animals, and
methods of transportation (see Figure\ \@ref(fig:E2-example-trial)). Different images were used in each block; there were 48 unique images total across the experiment.

@@

Stimulus sets were created to present participants with 24 presentation trials and 24 test trials on Day 1, and 48 recognition trials on Day 2. Each presentation trial screen presented two identical images (e.g. red apple and red apple). The test trial screens presented each of the previously-seen ("old") images paired with a new image (e.g. red apple and bicycle key). The recognition trial screens presented a single image that was either from the original set of 24 (seen twice across the first two phases) or was a corresponding foil for the original set (e.g. red apple or green apple). Thus, each participant was exposed to 72 unique images over the course of the experiment (the 24 foils for the "new" images in the test phase are never seen). Two full stimulus lists were created to counterbalance images across participants; the images that composed the "old" set for one list made up the "new" set for the list. The experimental design is visually depicted in Figure\ \@ref(fig:E3-design-schematic).

@@

During a trial, four pictures were displayed (target animal, target
instrument, distractor animal, distractor instrument), one in each
corner of the screen, and participants heard an audio prompt that
contained instructions about the action they needed to act out (e.g.,
"Rub the butterfly with the crayon"; see Figure\ \@ref(fig:E4-example-trial))[^2]. Using their
cursor, participants could act out the instructions by clicking on
objects and moving them or motioning over the objects[^3]. After the
action was completed, the participants were instructed to press the
space bar which led to a screen that said "Click Here" in the middle in
order to remove bias in the eye and mouse movements from the previous
trial. The experiment only allowed the participants to move on to the
next trial once the audio was completely done playing and the mouse had
been moved over at least one object.

[^1]: The WebGazer.js patch can be found at https://github.com/brownhci/WebGazer/commit/35301e56c192ef265bb270873a6526205f183711.

@@

All stimuli, experiment scripts, data, and analysis scripts are
available on the Open Science Framework at [https://osf.io/eubsc/](https://osf.io/eubsc/). 
The study pre-registration is available at
[https://osf.io/tv57s](https://osf.io/tv57s).

@@

During each trial of the main task, two faces were displayed on the two
halves of the screen, one on the left and one on the right (as in Figure\ \@ref(fig:E5-example-trial)). In the attractiveness task,
participants were asked to chose the more attractice face in the pair
and in the shape judgment task participants were asked to pick the face
that appeared rounder. They pressed the "a" key on their keyboard to
select the face on the left and the "d" key to select the face on the
right. A fixation cross appeared in the center of the screen between
each set of faces. Participants were asked to look at this fixation
cross in order to reset their gaze in between trials. The order of
the 19 face pairs was random for each participant.

@@

## Results
In the original study, a video-based eye tracker was used. The eye
movements of participants were recorded with a digital camera
downsampled to 33.3 Hz, with eye position determined
automatically with MediaAnalyzer software. In our study, subjects
supplied their own cameras, so hardware sampling rate varied. However,
data was collected at 20 Hz. <!--`r colorize("[TODO - CONFIRM]", 'red')` AJ: I don't know how to confirm this! -->

@@

 -->
```

# General Discussion

@@

Note also that while ROIs the size of quadrants are feasible, good calibration is important. Based on the results above, researchers using such paradigms may want to exclude data from participants with less than 50% validation accuracy. While stricter thresholds may decrease noise, they come at a cost. In preliminary analysis, we found that a threshold of 75% somewhat increased effect size in Exp. 4, but at the cost of reducing the sample size from 35 to 19. However, it could be helpful to more systematically investigate the optimal threshold and whether this varies by size of ROI.

There are a number of other interventions that might increase precision and should be investigated. For instance, it is possible to monitor head distance in real time [for an example, see @jiang2021easyeyes]; doing so and providing real-time feedback to participants should diminish noise. It may also be possible to automatically detect challenges to calibration accuracy (such as poor lighting) and provide targeted feedback. Similarly, it may be profitable to systematically study the parameters of eyegaze validation, for instance the number of validation points and the maximum distance allowed between gaze and target (set to 200px in the present study). These and other options should be investigated in order to determine how much they might improve precision.

Second, researchers need to take care with timing accuracy. Running studies on modern computers requires synchronizing multiple clocks, something that does not happen automatically. Failure to synchronize can result in substantial inaccuracies [@slim2022moving]. As indicated by the current work, the current jsPsych implementation of WebGazer.js has timing synchronization sufficient for many studies. We urge researchers to avoid using the current release of WebGazer.js as is, and instead either use jsPsych, our patch to WebGazer.js (https://github.com/brownhci/WebGazer/commit/35301e56c192ef265bb270873a6526205f183711), or other software that has been specifically documented as addressing common issues in timing for web-based eye-tracking.

Third, it is true for any eyetracker that precise data depends on good calibration, and good calibration depends in part on lighting and background [see, for instance, @yuksel2023investigation]. We suspect that it is possible to give subjects automated feedback on how to improve their own calibration, which could result in higher-quality data and fewer subject exclusions based on poor calibration.

@@

# Contributions
* Joshua R. de Leeuw
    + Conceptualization
    + Data curation
    + Formal analysis
    + Investigation
    + Methodology
    + Project administration
    + Software
    + Supervision
    + Validation
    + Visualization
    + Writing - original draft
    + Writing - review & editing

* Ariel N. James, Rachel Ryskin, and Joshua K. Hartshorne
    + Conceptualization
    + Formal analysis
    + Visualization
    + Writing - original draft
    + Writing - review & editing

* Undergraduate student authors, listed alphabetically: Haylee Backs, Nandeeta Bala, Laila Barcenas-Meade, Samata Bhattarai, Tessa Charles, Gerasimos Copoulos, Claire Coss, Alexander Eisert, Elena Furuhashi, Keara Ginell, Anna Guttman-McCabe, Emma (Chaz) Harrison, Laura Hoban, William A. Hwang, Claire Iannetta, Kristen M. Koenig, Chauncey Lo, Victoria Palone, Gina Pepitone, Margaret Ritzau, Yi Hua Sung, and Lauren Thompson
    + Investigation
    + Methodology
    + Software

@@

# Funding information
Funding for data collection was provided by the Vassar College Department of Cognitive Science (JdL). Partial funding for JKH's contributions was provided by NSF2436092.

@@

# Competing interests
The authors have no competing interests to declare.

@@

# Data accessibility statement
All data and analysis code underlying this paper can be found in the Github repository: https://github.com/jodeleeuw/219-2021-eyetracking-analysis. Study materials and pre-registrations for each experiment are available on Open Science Framework:

* Experiments 1a and 1b: [https://osf.io/g6dum/]
* Experiment 2: [https://osf.io/xezfu/]
* Experiment 3: [https://osf.io/k63b9/]
* Experiment 4: [https://osf.io/x3c49/]
* Experiment 5: [https://osf.io/eubsc/]
